---
title: "HW1: Data Science Essentials and Basic R Code"
format: html
editor: visual
---

# Section One

## Questions

### What do you think being a data scientist is about?

I think being a data scientist is about analyzing and understanding the large amounts of data that are available. While the data may not have been collected in a systematic way to test a given hypothesis, data scientists are able to engineer and format potentially messy and unstructured data in order to perform both simple exploratory analyses or use the data in a model.  

### What differences/similarities do you see between data scientists and statisticians?

Similarities:

-  Clean data is essential. No statistics or data science question can be answered if inappropriate, inaccurate, or insufficient data are collected. Understanding the context of the data is also essential.

-  Communication and proper visualization are huge parts of both jobs to ensure that the results can be properly articulated and explained in context. 

-  Both the term "statistician" and "data scientist" can be used to describe a diverse variety of roles that and there are many different tools available to both roles. Both jobs likely involve some style of coding and some style of analysis. 

Differences:

-  Statisticians can be more theoretical than data scientists. For instance, statisticians tend to be interested in developing and testing specific hypotheses, while data scientists focus on assessing correlation and conducting other simple exploratory analyses on existing data. 

-  Similarly, statisticians tend to play a more active role in designing where their data come from. They do this by designing surveys and other data collection instruments, then sometimes manually collecting responses. This is different from data scientists who use data produced automatically during a company's processes. This relates to the first bullet because more particular data/conducting a controlled experiment can be necessary to determine causation and evaluate a hypothesis, while data scientists can fit any model to their data after it has been collected. 

-  Due to the differences mentioned above, statisticians and data scientists can tend to use different tools. Statisticians are more likely to only use R, Python, and SQL. Data scientists may be expected to also have data engineering skills like Spark or Hadoop. Data scientists are also more likely to use machine learning techniques.

-  While clean data is needed for both roles, the process of data cleaning may be more involved for data scientists than statisticians. 


### How do you view yourself in relation to these two areas?

This is a great question and definitely something I have been trying to figure out! I worked full-time as a data _analyst_ for two years after graduating from college, and I found that I actually really enjoyed the more simple data cleaning parts of that job. I was very happy simply creating datasets from various file sources and then exploring the data on a surface level. In my classes at NC State, I have done better and been more motivated in my programming and applied courses compared to theoretical ones. As a result I have been wondering if more data science-y roles would be better suited to me than statistician roles. However, I don't have much experience with machine learning and algorithm techniques. Additionally, I am really interested in human/health research and feel that the research design aspect of statistics can lead to interesting questions/interventions. I think as I grow my confidence in statistical theory and design, I will maybe start to enjoy that aspect more. 

# Section Two

```{r}
y <- density(iris$Petal.Width)
```

The R code above saves the kernel density estimates of the variable Petal Width from the iris dataset in the R object "y". 

```{r}
class(y)
typeof(y)
str(y)
```

The R code above determines the class, type, and structure of the R object "y".

```{r}
#| echo: FALSE
plot(y)
```

The R code above plots the R object "y". 